Vragenlijst thesis evaluatie

### 2015-08-05 doc/template/chap-5.tex
verbeteringen jeroen

An Advanced Persistent Threat (referred to for the remainder of this work as APT) is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time. The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks. APT's are `advanced' because these attacks are well funded and (usually) the attacker needs a great amount of expertise to successfully penetrate a network. Not all APT's are \textit{technically} advanced though. The attacker can also try to exploit known vulnerabilities, in the hope that his target has not yet secured himself against them. \\
`Persistent' refers to the fact that the attacker never quits trying to attack his target. The attack can be spread over several years, taking multiple steps.  \\
An APT can be a mix of different types of malware and may use various propagation methods like the way a virus, worm or trojan horse propagates. \\
%
%Some examples of the biggest most rare APTs are listed to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}] 
%
%\subsubsection{Equation}
%Equation is a complex cyberattack platform where the first known sample is from 2002, but it was only discovered 12 years later in 2014. This APT propagates through usb drives, cd or physical media. It will search for exploits and will self-replicate itself to spread the infection. The purpose of this virus is to steal data and cyberespionage. 
%
%\subsubsection{Regin}
%
%\subsubsection{Flame}
%Way of propagation through USB drives, LAN spreading. purpose cyber espionage.
%
%\subsubsection{Black energy}
% purpose cyber espionage and DDoS, data wiping. prop usb lan

According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed over 2.54 million dollar. Companies need a defence mechanism to defend themselves against APT's. As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified. To mitigate these kind of attacks another security countermeasure is needed. \\
%This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
In this paper we will propose an appropriate game-theoretical modelling of defending a network against APTs and analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.

\section{A brief introduction in Game Theory}
\label{Cha1:briefintro}
Gametheory is a mathematical study to analyse interactions between independent and self-interested agents. To get an understanding of the most important concepts of game theory, a short introduction based on the work of 
\cite{leyton2008essentials} and \cite{Coursera} is given in this section. For a more detailed and full introduction to game theory, the reader is referred to 
\cite{leyton2008essentials}.  \\
%In section \ref{cha1:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
%The last section \ref{ch1:extendedWork} will cover the extensions and additions already made on FlipIt. \\
%------------------------------------------------%
%            Intro Game Theory 					 %
%------------------------------------------------%
Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes. Game theory has therefore important applications in many area's such as economics, politics, biology, computer science, philosophy and a variety of other disciplines. It has come only in ascent during the Second World War with Oskar Morgenstern and John von Neumann both publishing a book on game theory, entitled "Theory of Games and Economic Behavior" (Game theory and economic behaviour) in 1944. This book was about the mathematical analysis of a series of thinking games. The book described a series of games making a distinction between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies (e.g. chess) and games in which this has an influence (e.g. poker). John Nash also played a major role in the history of game theory. He was one of the mathematicians who has formalized game theory. The Nash equilibrium was named after him.
~~\\
%Every agent has different levels of happiness for the different outcomes.
%self interested meaning

One of the assumptions underlying game theory is that the players of the game, the agents, are independent and self-interested. This does not necessarily mean that they want to harm other agents or that they only care about themselves. 
%utility function meaning 
Instead it means that each agent has preferences about the states of the world he likes. These preferences are mapped to natural numbers and are called the utility function. The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the states of the world. \\
%Cooperative and non cooperative games
In a decision game-theoretic approach an agent will try to act in such a way to maximise his expected or average utility function. It becomes more complicated when two or more agents want to maximise their utility and when actions of the agents can affect each other's utilities. This kind of games are referred to as non-cooperative game theory, where the basic modelling unit is the group of agents. The individualistic approach, where the basic modelling is only one agent, is referred to as cooperative game theory.

@@

%John Nash speelde ook een grote rol in de geschiedenis van de speltheorie. Hij is een van de wiskundigen geweest die speltheorie geformaliseerd heeft. Het Nash evenwicht werd naar hem vernoemd. Een Nash evenwicht wordt gezien als een evenwicht tussen beide spelers zodat ze allebei de beste tactiek kiezen en niet meer veranderen als de andere van tactiek veranderen. John Nash breide de theorie over het Nash evenwicht in een paper nog uit met gemengde strategieën. In 1994 kreeg John Nash samen met twee andere wiskundigen gespecialiseerd op het vlak van speltheorie de Nobelprijs voor de economie op basis van hun prestaties in de niet-coöperatieve speltheorie. . Over John Nash is een prachtige film 
\subsubsection{Best response and Nash Equilibrium}
One of the solution concepts in Game Theory for non-cooperative games that will be used in this paper is a Nash Equilibrium. A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game. To define this concept we first introduce the concept of best response. The best response for a player is the action of a player that maximizes it's pay-off for any given action of the other player. We define $Opt_{i}$ as the best response function for player \textit{i}. The best response for player \textit{1} is: $a_{1} = Opt_{1}(a_{2})$, given that $a_{2}$ is an action of player 2. A 
For a Nash Equilibrium each player has a list of actions and each player's action maximizes his or her pay-off given the actions of the other players. Nobody has the incentive to change his or her action if an equilibrium profile is played. We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$.\\

An example to explain a Nash equilibrium of a two-player game is the prison dilemma. In this game there are two players who are both rational and both of them have committed a crime. `Rational' means that they want the best for themselves and it is not their purpose to do harm to others. Both are locked in a separate room and they cannot tell in advance to each other what they are going to say. Each of them can betray the other or they can support each other and remain silent. If a player confesses, he will, depending on what the other does, go for three years in prison or he is free to go. If the player is silent, he will, depending on what the other player does, go for five years in jail or only one year. \\

@@

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.6]{Images/PrisDilemma}
\caption{Prisoners dilemma: An example from the field of Game Theory. It is a game between two perfectly rational prisoners who don not know what the other one will do. Each can confess or stay silent. The resulting sentence is shown below each prisoner. }
\label{Prison}
\end{figure}

@@

The game begins at $t=0$ and continues indefinitely ($t \rightarrow \infty $). The time in the game is assumed to be continuous. To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time. 
%A flip will be regarded as a move from a player \textit{i}. 
Each move implies a certain cost $k_{i}$ and can vary for each player. Both players try to minimize their cost. Adding a cost prevents players to move too frequently. \\

The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player has no clue whether the other player (his adversary) has flipped the resource or not. For instance, the defender does not know if the resource has been compromised by the attacker until he flips the resource himself. The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves. A move can also result in a "wasted move", called a flop. It may happen that the resource was already under control of the player. If the player moves when he or she has already control over the resource, he or she would have wasted a move since this does not result in a change of ownership, so the cost is wasted. \\

@@

The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}. It is expressed as follows:
\begin{equation}\label{first}
G_{i}(t) = \int_0^t \! C_{i}(x) dx.
\end{equation}
If we add up the gain of the defender and the gain of the attacker it should sum up to \textit{t}:
\begin{equation}\label{first}
G_{D}(t) + G_{A}(t) = t
\end{equation}
The average gain rate of player \textit{i} is defined as:
\begin{equation}\label{first}
\gamma_{i}(t) = G_{i}(t)/t.
\end{equation}
And thus for all $t > 0$ :
\begin{equation}\label{first}
\gamma_{D}(t) + \gamma_{A}(t) = 1
\end{equation}
~$ \alpha_{i}$ Defines the average move rate by player \textit{i} up to time \textit{t}.
Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
\begin{equation}\label{first}
\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
\end{equation}
This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time\textit{ t} will increase to infinity and the average benefit may not have limiting values.
\begin{equation}
\beta_{i}(t)  = \lim_{t \to \infty} inf \beta_{i}(t) 
\end{equation}
\\

\subsubsection{Strategies}
Because the players move in a stealthy way, there are different types of feedback that a player can get while moving. These types of feedback can be divided into two groups of strategies. The non-adaptive strategies and the adaptive strategies. These are described in table \ref{table:Strategies}.\\

If there is no feedback for either player, we have a non-adaptive strategy. Because a player does not receive any feedback during the game he will play in the same manner against every opponent. The strategy is called non-adaptive because the playing strategy is not dependent on the opponents movements. An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process. An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval. An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed. \\

In case there is feedback, a player can adapt his strategy to the information received about the opponent's moves. Depending on the amount of information received, two subclasses of adaptive strategies can be identified. The Last Move (LM) strategies represent the class where whenever a player flips he will find out the exact moment that the opponent moved the last time. In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's moves. \\

@@

\subsubsection{Results of the FlipIt game}
The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results:  
\begin{itemize}
\item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy. This means that it is always advantageous for the opponent to play periodically against a player with a non-adaptive strategy;
\item periodic games are disadvantageous against players following a Last Move adaptive strategy;
\item if a defender facing an LM attacker and with a cost much lower than the attacker, he can play with a periodic rate that is fast enough he'll force the attacker to drop out;
\item a defender facing an LM attacker has to play with a randomized strategy e.g. exponential strategy;
\item any amount of feedback about the opponent received during the game, benefits to a player.
\end{itemize}

@@

Various possible ways to extend FlipIt have already been proposed. 
Laszka et al. made a lot of additions and extensions to the original game of FlipIt. For instance Laszka et al. extended the basic FlipIt game to multiple resources. The rationale is that for compromising a system in real life, more than just one resource needs to be taken over. An example is that gaining access to deeper layers of a system may require breaking several passwords. The model is called FlipThem \cite{FlipThem}. Laszka et al. also use two ways to flip the multiple resources: the AND and the OR control model. In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system. \\

Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
is extending the game to also consider non-targeted attacks by non-strategic players. In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks. Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities. \\
One of the last important additions from Laszka et al. \cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately. This approach is similar to what will be done in this paper, but nevertheless has some major differences. In Laszka's paper, firstly the moves by the attacker are still covert but the moves made by the defender are known to the attacker. This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender. Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT. Some ATP's are designed only to cause harm on the systems that are infected e.g. the Stuxnetworm that was developed to destroy nuclear reactors. The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet. Another example is by the use of honeypots. Honeypots emulate services or create multiple instances of real operating systems and can pretend to have sensitive information.  Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback. This is a way how the defender can also be stealthy by providing  false information through the honeypot. The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently. For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate. The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process. In our paper the delay is given by one parameter, that can be the result of any virus propagation model. The third and last difference is that the paper of Laska has multiple attackers and they try to find the best strategy of the defender against both targeted and non-targeted attacks. The conclusion of this paper is that the optimal strategy for the defender is moving periodically. \\

FlipIt also has been applied to several cases in computer security. Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}. \\
Other authors used the FlipIt game to apply it to a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker. This could be verified by the defender, who has an extra move ``test'' beside the flip move. The basic idea is to test with an extra action if the resource has been compromised or not. This move also involves an extra cost.\\
A three-player game has also been investigated where the FlipIt framework of two players is extended by another player. This player represents an insider that trades value information with the attacker \cite{fengstealthy}.\\

Finally researchers also have investigated the behaviour of humans playing FlipIt. A. Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions. They found out that the results improve over time but that they are dependent on gender, age, and other individual difference variables. The result also shows that the participants perform generally better when they have more information about the strategy of the opponent, which is a computerized player. Reitter et al. \cite{reitter2013risk} extended the work of A. Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.\\

>>>>

An Advanced Persistent Threat (referred to for the remainder of this work as APT) is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time. The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks. APT's are `advanced' because these attacks are well funded and because the attacker (usually) needs a great amount of expertise to successfully penetrate a network. Not all APT's are technically advanced though. The attacker can also try to exploit known vulnerabilities, in the hope that his target has not yet secured itself against them. \\
`Persistent' refers to the fact that the attacker isn't trying to gain immediate results, and the attacker won't stop trying after one failed attempt. The attack can be spread over several years, taking multiple steps.  \\
An APT can be a mix of different types of malware and may use various propagation methods. \\
% like the way a virus, worm or trojan horse propagates. \\
%
%Some examples of the biggest most rare APTs are listed to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}] 
%
%\subsubsection{Equation}
%Equation is a complex cyberattack platform where the first known sample is from 2002, but it was only discovered 12 years later in 2014. This APT propagates through usb drives, cd or physical media. It will search for exploits and will self-replicate itself to spread the infection. The purpose of this virus is to steal data and cyberespionage. 
%
%\subsubsection{Regin}
%
%\subsubsection{Flame}
%Way of propagation through USB drives, LAN spreading. purpose cyber espionage.
%
%\subsubsection{Black energy}
% purpose cyber espionage and DDoS, data wiping. prop usb lan

According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed 2.5 million dollars. As such, companies need a defence mechanism to defend themselves against APT's. As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified. To mitigate these kind of attacks another security countermeasure is needed. \\
%This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
This paper proposes an appropriate game-theoretical modelling of defending a network against APTs and will analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.

\section{A Brief Introduction in Game Theory}
\label{Cha1:briefintro}
Game theory is a mathematical study to analyse interactions between independent and self-interested agents. To get an understanding of the most important concepts of game theory, a short introduction based on the work of 
\cite{leyton2008essentials} and \cite{Coursera} is given in this section. For a more detailed and complete introduction to game theory, the reader is referred to 
\cite{leyton2008essentials}.  \\
%In section \ref{cha1:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
%The last section \ref{ch1:extendedWork} will cover the extensions and additions already made on FlipIt. \\
%------------------------------------------------%
%            Intro Game Theory 					 %
%------------------------------------------------%
Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what each agent does and how these interactions should be structured to lead to good outcomes. Game theory therefore has important applications in many areas such as economics, politics, biology, computer science, philosophy and a variety of other disciplines. It gained recognition during the Second World War, when Oskar Morgenstern and John von Neumann both published a book on game theory, titled "Theory of Games and Economic Behavior" in 1944 \cite{von1944theory}. This book addressed the mathematical analysis of a series of thinking games. A distinction was made between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies (e.g. chess) and games wherein this factor does have an influence (e.g. poker). John Nash also played a major role in the history of game theory. He was one of the mathematicians who has formalized game theory. The Nash equilibrium, a common solution concept of a non-cooperative game, was named after him.
~~\\
%Every agent has different levels of happiness for the different outcomes.
%self interested meaning

One of the underlying assumptions game theory is that the players of the game are independent and self-interested. This does not necessarily mean that they want to harm other agents or that they only care about themselves. 
%utility function meaning 
Instead it means that each agent has preferences about which states of the world he likes. These preferences are mapped to natural numbers and are called the utility functions. The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the states of the world. \\
%Cooperative and non cooperative games
In a decision game-theoretic approach an agent will try to act in such a way as to maximise his expected or average utility function. It becomes more complicated when two or more agents want to maximise their utility and when actions of an agent can affect other agents' utilities. In this kind of game, the basic modelling unit is the group of agents. It is referred to as \textit{non-cooperative game theory}. In the individualistic approach the basic modelling is only one agent. This type of game is referred to as \textit{cooperative game theory}.

@@

%John Nash speelde ook een grote rol in de geschiedenis van de speltheorie. Hij is een van de wiskundigen geweest die speltheorie geformaliseerd heeft. Het Nash evenwicht werd naar hem vernoemd. Een Nash evenwicht wordt gezien als een evenwicht tussen beide spelers zodat ze allebei de beste tactiek kiezen en niet meer veranderen als de andere van tactiek veranderen. John Nash breide de theorie over het Nash evenwicht in een paper nog uit met gemengde strategieën. In 1994 kreeg John Nash samen met twee andere wiskundigen gespecialiseerd op het vlak van speltheorie de Nobelprijs voor de economie op basis van hun prestaties in de niet-coöperatieve speltheorie. . Over John Nash is een prachtige film 
\subsubsection{Best response and Nash Equilibrium}
One of the solution concepts in Game Theory for non-cooperative games that will be used in this paper is a Nash Equilibrium. A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game. To define this concept we first introduce the concept of best response. The best response for a player is the action of a player that maximizes it's pay-off for any given action of the other player. We define $Opt_{i}$ as the best response function for player \textit{i}. The best response for player \textit{1} is: $a_{1} = Opt_{1}(a_{2})$, given that $a_{2}$ is an action of player 2. For a Nash Equilibrium each player has a list of actions and each player's action maximizes his or her pay-off given the actions of the other players. Nobody has the incentive to change his or her action if an equilibrium profile is played. We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$.\\

An example to explain a Nash equilibrium of a two-player game is the prison dilemma. In this game there are two players who are both rational and both of them have committed a crime. `Rational' means that they want the best for themselves and it is not their purpose to do harm to others. Both are locked in a separate room and they cannot tell in advance to each other what they are going to say. Each of them can betray the other or they can support each other and remain silent. If a player confesses, he will either be imprisoned for 3 years or go free, depending on what the other does. If the player is silent, he will either be imprisoned for 5 years or one year, depending on what the other does.  \\ \todo{nakijken met praten, bekennen of andere veraden?}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.6]{Images/PrisDilemma}
\caption{Prisoners dilemma: An example from the field of Game Theory. It is a game between two perfectly rational prisoners who do not know what the other one will do. Each prisoner can confess or stay silent. The resulting sentence is shown below each prisoner. }
\label{Prison}
\end{figure}

@@

The game begins at $t=0$ and continues indefinitely ($t \rightarrow \infty $). The time in the game is assumed to be continuous. To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time. 
%A flip will be regarded as a move from a player \textit{i}. 
Each move implies a certain cost $k_{i}$ and can vary for each player. Both players try to minimize their cost. Adding a cost prevents players from moving too frequently. \\

The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player does not immediately finds out if the other player (his adversary) has flipped the resource or not. A player only finds out about the state of the game when he moves himself. The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves. A move can also result in a "wasted move", called a flop. It may happen that the resource was already under control of the player. If the player moves when he or she has already control over the resource, he or she would have wasted a move since this does not result in a change of ownership, so the cost is wasted. \\

@@

The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost for player \textit{i} is denoted by $k_{i}$. 
The total gain for player \textit{i} is equal to the total amount of time that player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}. It is expressed as follows:
\begin{equation}\label{first}
G_{i}(t) = \int_0^t \! C_{i}(x) dx.
\end{equation}
The gain of the attacker and the defender always sums up to \textit{t}:
\begin{equation}\label{first}
G_{D}(t) + G_{A}(t) = t
\end{equation}
The average gain rate of player \textit{i} is defined as:
\begin{equation}\label{first}
\gamma_{i}(t) = G_{i}(t)/t.
\end{equation}
And thus for all $t > 0$ :
\begin{equation}\label{first}
\gamma_{D}(t) + \gamma_{A}(t) = 1
\end{equation}
~Let $ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t} and let $\beta_{i}(t)$ denote player's \textit{i} average benefit rate upto time \textit{t}:
\begin{equation}\label{first}
\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
\end{equation}
This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time\textit{ t} will increase to infinity and the average benefit may not have limiting values.
\begin{equation}
\beta_{i}(t)  = \lim_{t \to \infty} inf \beta_{i}(t) 
\end{equation}
\\

\subsubsection{Strategies}
Because the players move in a stealthy way, there are different types of feedback a player can get by flipping the resource. These types of feedback can be divided into two groups of strategies. The non-adaptive strategies and the adaptive strategies. These are described in table \ref{table:Strategies}.\\

If there is no feedback for either player, he will play in the same manner against every opponent. The strategy is then called \textit{non-adaptive} because the playing strategy is not dependent on the opponents movements. An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process \todo{renewal beter uitleggen}. An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval. An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed. \\

If the players receive feedback, a player can adapt his strategy to the information received about the opponent's moves. This strategy is called \textit{adaptive} strategies. Depending on the amount of information received, two subclasses of adaptive strategies can be identified. The Last Move (LM) strategies represent the class where, whenever a player flips, he will find out the exact moment that the opponent moved the last time. In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's moves. \\

@@

\subsubsection{Results of the FlipIt game}
The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results \cite{FlipIt}:  
\begin{itemize}
\item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy. This means that it is always advantageous for the opponent to play periodically against a player with a non-adaptive strategy;
\item periodic games are disadvantageous against players following a Last Move adaptive strategy;
\item if a defender facing an LM attacker and with a cost much lower than the attacker, he can play with a periodic rate that is fast enough he'll force the attacker to drop out;
\item a defender facing an LM attacker has to play with a randomized strategy e.g. exponential strategy;
\item any amount of feedback about the opponent received during the game, benefits to a player.
\end{itemize}

@@

Various possible ways to extend FlipIt have already been proposed. 
Laszka et al. made a lot of additions and extensions to the original game of FlipIt. For instance Laszka et al. extended the basic FlipIt game to multiple resources. The rationale is that for compromising a system in real life, more than just one resource needs to be taken over. For example, gaining access to deeper layers of a system may require breaking several passwords. This model is called FlipThem \cite{FlipThem}. Laszka et al. also use two ways to flip the multiple resources: the AND and the OR control model. In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system. \\

Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
is extending the game to also consider non-targeted attacks by non-strategic players. In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks. Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities. \\
One of the last important additions from Laszka et al. \cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately. This approach is similar to what will be done in this paper, but nevertheless has some major differences. Firstly, in Laszka's paper, the moves by the attacker are still covert but the moves made by the defender are known to the attacker. This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender. Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT. Some ATP's are designed only to cause harm on the systems that are infected e.g. the Stuxnetworm that was developed to destroy nuclear reactors. The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet. Another example is the use of honeypots. Honeypots emulate services or create multiple instances of real operating systems and can pretend to have sensitive information.  Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback. By providing false information through honeypots, the defender can also remain stealthy. The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently. For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate. The non-targeted attacks are modelled as a single attacker and the time until it succeeds is given by a Poisson process. In this paper the delay is given by one parameter, which can be the result of any virus propagation model. The third and last difference is that the paper of Laska has multiple attackers who try to find the best strategy of the defender against both targeted and non-targeted attacks. The conclusion of this paper is that the optimal strategy for the defender is moving periodically. \\

FlipIt also has been applied to several cases in computer security. Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}. \\
Other authors used the FlipIt game to apply it to a specific scenario. To be able to use the FlipIt game, modifications were required for the FlipIt model.
One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker. This could be verified by the defender, who has an extra move ``test'' beside the flip move. The basic idea is to test with an extra action if the resource has been compromised or not. This move also involves an extra cost.\\
A three-player game has also been investigated where the FlipIt framework of two players is extended by another player. This player represents an insider that trades value information with the attacker \cite{fengstealthy}.\\

Finally, researchers have also investigated the behaviour of humans playing FlipIt. A. Nochenson and Grossklags \cite{nochenson2013behavioral}  investigated how people really act when given temporal decisions. They found out that the results improve over time but that they are dependent on gender, age, and other individual difference variables. The result also shows that the participants perform generally better when they have more information about the strategy of the opponent, which is a computerized player. Reitter et al. \cite{reitter2013risk} extended the work of A. Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.\\
